\chapter{Conclusions}

This thesis addressed the problem of file synchronization for small collaborative teams that need something simpler than Git but safer than basic file mirroring.

We identified a gap in existing tools. Centralized services need external infrastructure and accounts. Version control systems require technical knowledge and manual workflows. Peer-to-peer sync tools lack version safety. No existing tool combines automatic local-network synchronization with built-in version backups and zero configuration.

Our system fills this gap with a two-process architecture. The Go networking daemon handles peer discovery (mDNS/DNS-SD), synchronization coordination (Lamport clocks with LWW conflict resolution), and TCP file transfer. The C++ storage daemon handles file watching, hashing, backups, and disk I/O. Together, they provide a shared directory that behaves like a normal folder while synchronizing across peers on the same network.

The design makes trade-offs that fit our target use case:

\begin{itemize}
    \item Lamport clocks instead of vector clocks: simpler but cannot detect truly concurrent edits.
    \item Last-Write-Wins instead of CRDTs: simpler but one version is discarded (with backup).
    \item Whole-file transfers instead of delta sync: simpler but uses more bandwidth.
\end{itemize}

These choices work for 3 to 5 peers on a local network. Scaling further would require revisiting them.

Future work includes:

\begin{itemize}
    \item Block-level delta synchronization using the rsync rolling-checksum algorithm to reduce bandwidth for large files.
    \item CRDT-based metadata merging to allow conflict-free convergence without LWW.
    \item Cross-subnet discovery for peers on different networks.
    \item Mobile platform support, where filesystem access and background execution work differently.
\end{itemize}

The project shows that a simple, serverless synchronization tool is feasible by combining well-understood protocols (mDNS, TCP, SHA-256) with a practical conflict resolution approach (LWW with backups).
